---
title: 'Causal Inference Models: Asthma and Traffic Pollution'
output:
  html_document: default
  html_notebook: default
---

General purpose of this project is to try and extend causal inference methods, particularly the counterfactual/potential outcomes framework to non-binary exposures (continous, categorical).

Try out the 'ipw' package.

## Replication of Sheryl's Dissertation Results

Importing SAS files of Oakland Kicks Asthma data provided by Sheryl.

```{r import and data wrangle}
library(haven) # library for importing SAS files
library(tidyverse) # data wrangle package

no2layer <- read_sas("./no2layer.sas7bdat") %>% # summary of NO2
  select(OKAID, NO2)
  
no2_col_names <- colnames(no2layer) %>% tolower() # lowercase all col names
colnames(no2layer) <- no2_col_names # assign lowercase col names to df

#head(no2layer) # check out the first 6 lines

okawork <- read_sas("./okawork1.sas7bdat")
#glimpse(okawork) # 433 variables; not priting

# join okawork with no2layer by okaid
asthma_df <- okawork %>% right_join(no2layer, by = "okaid") %>% 
  # create ed visit variable from sh_11 var (assuming 1=event)
  mutate(edvisit = ifelse(sh_11 == 1, 1,
                   ifelse(sh_11 == 2, 0, NA))) 


#head(asthma_df)
# sort column names
col_names_sorted <- colnames(asthma_df[, order(names(asthma_df))])
head(col_names_sorted)
summary(as.factor(asthma_df$sh_11))

```

### Descriptive Statistics

Compared to Sheryl's table 11, my numbers are smaller.

```{r descriptive}

# Table packages don't seem to be working properly. May need to update something
# load table library
# library(stargazer) # load table package

xtabs(~asthma + sh_11, asthma_df)
tab <- xtabs(~asthma + inbuff, asthma_df)
tab
prop.table(tab)
print(tab)

print(summary(tab))

# print(stargazer(tab, type = "html"))
```

```{r standard models, message=F}
# load broom package to get out tidy dataframes from models
library(broom)
# basic logic model
mod1 <- tidy(glm(asthma ~ inbuff, data = asthma_df, 
                 family = "binomial"(link = "logit")))
mod1


# check ed visits
xtabs(~edvisit, data = asthma_df)

# asthma ed visit mod
# crude 
ed_crude <- tidy(glm(edvisit ~ inbuff, data = asthma_df,
                    family = "binomial"(link = "logit")))

ed_crude



set.seed(777) # set seed
# try using bootstrap function in broom
ed_boot <- asthma_df %>% bootstrap(1000) %>% 
  do(tidy(glm(edvisit ~ inbuff, ., family = "binomial"(link = "logit"))))

# find CI from bootstrap
alpha = 0.05

bounds <- ed_boot %>% group_by(term) %>% 
  summarise(median = quantile(estimate, 0.5), 
            low = quantile(estimate, alpha/2),
            high = quantile(estimate, 1- alpha/2))

bounds



```

```{r marginal structural model}

# Marginal structural model approach ----

# Create unexposed and exposed dataframes

exp_data <- unexp_data <- asthma_df %>% 
  select(asthma, edvisit, inbuff, no2)

exp_data$inbuff <- 1
unexp_data$inbuff <- 0

# run standard model again of inbufrf
mod1 <- glm(edvisit ~ inbuff, data = asthma_df, 
                 family = "binomial"(link = "logit"))
summary(mod1)

# updated predicted probability in df where everyone is exposed
pred_outcome_exp <- predict(mod1, newdata=exp_data, type='response')
head(pred_outcome_exp)

# updated predicted probability in df where no one is exposed
pred_outcome_unexp <- predict(mod1, newdata=unexp_data, type='response')
head(pred_outcome_unexp)

# g-computation estimate (simple substitutions method)
mean(pred_outcome_exp)
mean(pred_outcome_unexp)

simple_sub <- mean(pred_outcome_exp - pred_outcome_unexp)
simple_sub


# try to now bootstrap the simple sub method ----
ed_vis_inbuff <- asthma_df %>% 
  select(edvisit, inbuff)

# create custom msm function ----
# using bootstrap funciton in broom, i only need to define dataframe
# as bootstrap feeds in the randomly sample /w replacement df to use
# I think eventually I'd like to work in arguments for outcome,
# exposure
msm.fun <- function(data, y, a, w){ 

  # if no covariate vector supplied
  if(missing(w)){ 
  col_n <- c(match(y, names(data)),
             match(a, names(data)))
  
  df_sample <- data %>% select(col_n)
  
  # create new a = 1, a = 0 datasets
  exp_data <- unexp_data <- df_sample
  
  # set all a vals in exposed dataframe to 1
  exp_data[,2] <- 1
  # set all a vals in unexp dataframe to 0
  unexp_data[,2] <- 0
  
  # run initial model on reduced df_sample 
  mod1 <- glm(as.formula(paste0(y, "~", a)), data = df_sample, 
          family = "binomial"(link = "logit"))  
  
  # else if covariates supplied  
  } else { # formula to run if covariates present
  # extract row numbers of y, a, and w vectors
  col_n <- c(match(y, names(data)),
             match(a, names(data)),
             match(w, names(data)))
  
  df_sample <- data %>% select(col_n)
  
  # create new a = 1, a = 0 datasets
  exp_data <- unexp_data <- df_sample
  
  # set all a vals in exposed dataframe to 1
  exp_data[,2] <- 1
  # set all a vals in unexp dataframe to 0
  unexp_data[,2] <- 0
  
  # run initial model on reduced df_sample 
  mod1 <- glm(as.formula(paste0(y, "~", a, "+",  
          paste(w, collapse = "+"))), data = df_sample, 
          family = "binomial"(link = "logit"))
  
  } # end formula with covariates
  
  
  # updated predicted probability in df where everyone is exposed
  pred_outcome_exp <- predict(mod1, newdata = exp_data, 
                              type='response')
  # updated predicted probability in df where no one is exposed
  pred_outcome_unexp <- predict(mod1, newdata = unexp_data,
                                type='response')
  # g-computation estimate (simple substitutions method)
  # risk difference
  risk_diff <- mean(pred_outcome_exp - pred_outcome_unexp)
  # risk ratio
  risk_ratio <- mean(pred_outcome_exp)/mean(pred_outcome_unexp) 
  
  # return values
  msm_est <- c(risk_diff, risk_ratio)
  names(msm_est) <- c("Risk Difference", "Risk Ratio")
  return(msm_est)
  
  } # end function


test <- msm.fun(asthma_df, y = "edvisit", a = "inbuff", w = c("male", "white"))
test

# trying boot with broom bootstrap function (need tidy funciton as well)
msm_boot <- asthma_df %>% 
  # filter to only variables i'm interested in
  # random sample with replacement boot
  bootstrap(1000) %>% 
  # do formula
  do(tidy(msm.fun(., y = "edvisit", a = "inbuff", w = c("male", "white")))) # call msm formula


# find CI from bootstrap
alpha <- 0.05

bounds <- msm_boot %>% group_by(names) %>% 
  summarise(median = quantile(x, 0.5), 
            low = quantile(x, alpha/2),
            high = quantile(x, 1- alpha/2))

bounds


# simple models ----
# risk diff
rd_mod <- tidy(glm(edvisit ~ inbuff + male + white, data = asthma_df, 
                 family="binomial"(link = "identity")))

rd_95ci <- rd_mod %>% filter(term == "inbuff") %>% 
  mutate(lower_95 = estimate - (1.96*std.error),
         upper_95 = estimate + (1.96*std.error)) %>% 
  select(estimate, lower_95, upper_95)

rd_mod
rd_95ci

normal_mod <- glm(edvisit ~ inbuff + male + white, data = asthma_df, 
                 family="gaussian") 
summary(normal_mod)

summary(glm(edvisit ~ white, data = asthma_df, family="poisson"(link = "log")))

# risk ratio
rr_mod <- tidy(glm(edvisit ~ inbuff + white, data = asthma_df, 
                 family="poisson"(link = "log")))

rr_95ci <- rr_mod %>% filter(term == "inbuff") %>% 
  mutate(lower_95 = estimate - (1.96*std.error),
         upper_95 = estimate + (1.96*std.error)) %>% 
  select(estimate, lower_95, upper_95)

exp(rr_95ci)


```

