---
title: "NO2 and asthma using TMLE"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---


## Introduction

NO2 is bad. We want to know if NO2 increases the likelihood of reporting an asthma ED visit in children with asthma living in Oakland. 

## Research Question

Overall Aim: Compare and contrast causal inference statistical methods to standard regression and their utility in environmental epidmeiology studies.

Standard Regression: Estimate the continous relationship between increasing yearly? average of NO~2~ and the likelihood of a reported ED visit in children with asthma. 

Causal Inference: Estimate the marginal effect of a defined binary exposure and the likelihood of a reported ED visit in children with asthma under the counterfactual where everyone is exposed (Y=a1) and no one is expose (Y=a0).

Study population: Cohort of school children with asthma living in Oakland, California.
Outcome: Emergency room visit.
Exposure: NO~2~ exposure.

### Setup

Loading libraries.

```{r packages, message=F, warning=F, echo=F}
library(tidyverse) # tidyverse 
library(glmnet) # lasso/elastic net
library(hdps) # high density propensity score algorithm
library(SuperLearner) # superlearner ensemble algorightm
library(tmle) # caret ML processing
library(ctmle) # collaborative targeted maximum likelihood
library(caret) # data prep for ml models

# markdown options
knitr::opts_chunk$set(fig.height = 6, fig.width = 8, fig.align = "center",
                      progress = F, message = F, warning = F, echo=F,
                      results = "asis")
```

Preparing Oakland Kicks Asthma dataframe and joining with NO~2~ data frame.

```{r read_data}
# loading no2 value
no2layer <- haven::read_sas("./no2layer.sas7bdat") %>% # summary of NO2
  select(OKAID, NO2) %>% 
  mutate(OKAID = as.character(OKAID))
  
no2_col_names <- colnames(no2layer) %>% tolower() # lowercase all col names
colnames(no2layer) <- no2_col_names # assign lowercase col names to df

# loading oakland working
okawork <- haven::read_sas("./okawork1.sas7bdat") %>% 
  # convert okaid to character
  mutate(okaid = as.character(okaid))

# join okawork with no2layer by okaid to create final dataframe
asthma_no2_df <- okawork %>% right_join(no2layer, by = "okaid") %>% 
  # create ed visit variable from sh_11 var (assuming 1=event)
  mutate(edvisit = ifelse(sh_11 == 1, 1,
                   ifelse(sh_11 == 2, 0, NA))) %>% 
  # filter out missing edvisits
  filter(!is.na(edvisit)) %>%
  filter(asthma == 1) %>% 
  filter(sexmiss == 0 & racemiss == 0 & homelmiss == 0) %>% 
  # create no2 distributions
  mutate(no2_10dec = ifelse(no2 >= 19.9, 1, 0),
         no2_1qt = ifelse(no2 <= 16.237, 1, 0),
         no2_2qt = ifelse(no2 > 16.237 & no2 <= 17.180, 1, 0),
         no2_3qt = ifelse(no2 > 17.180 & no2 <= 18.090, 1, 0),
         no2_4qt = ifelse(no2 >= 18.090, 1, 0),
         no2_median = ifelse(no2 >= 17.180, 1, 0),
         age = as.numeric(round((timentry - dob)/365,0)),
         # imputate age based on grade
         age = ifelse(age < 10 & grade == 6, 11,
                  ifelse(is.na(age) & grade == 8, 14, age)))
```

Set up custom color theme.

```{r color_theme}
ryan_theme <- theme(panel.background = element_rect(fill = "black", 
          colour = "black", size = 0.5, linetype = "solid"),
        panel.grid.major = element_line(size = 0.2, colour = "white"),
        panel.grid.minor = element_line(size = 0.1, color = "white", 
                                        linetype = "dotted"),
        plot.background = element_rect(fill = "#141e30", colour="#141e30",
                                       size = 0.5, linetype = "solid"),
        text = element_text(colour = "white"),
        axis.text = element_text(colour = "white"),
        strip.background = element_blank(),
        strip.text = element_text(colour="white"),
        legend.background = element_blank(),
        legend.key = element_blank())
```

## Study Population Description

The Oakland Kicks Asthma study is a school-based study with n=4017 children.

Our study population consists of 20% of these school-aged children (n=771) that have a diagnosis of asthma living in Oakland. Data were collected from (dates). 
Our primary outcome is reported period prevalent emergency department visits in the past year *(Sheryl is this correct?)*. We'll evaluate some basic demographic characteristics of this population by reported emergency department visit.

###Demographics by ED Visit 

```{r demo_tab}
# subset df to demographic variables
demo <- asthma_no2_df %>% 
  select(edvisit, age, grade, male, female, enghome, spanhome,
         chnshome,viethome, olanghome,api, afra, latino, white, natam, 
         mixed, othrace) %>% 
  # change some vars to factors
  mutate_at(vars(-age), as.factor) 

# colname demographic vector except edvisit
demo_vec <- colnames(demo)[c(-1,-2)]

# age distribution 
age_tab <- demo %>% 
  group_by(edvisit) %>% 
  summarise(mu_age = round(mean(age),1)) %>% 
  mutate(var = "age_yr_mu",
         edvisit = case_when(edvisit == 1 ~ 'yes_val',
                             edvisit == 0 ~ 'no_val')) %>% 
  spread(edvisit, mu_age)

# n
age_tab <- demo %>% 
  group_by(edvisit) %>% 
  summarise(n=n()) %>% 
  mutate(var = "age_yr_mu",
         edvisit = case_when(edvisit == 1 ~ 'ed_yes_n',
                             edvisit == 0 ~ 'ed_no_n')) %>% 
  spread(edvisit, n) %>% 
  left_join(age_tab, by="var")

# precent table
prec_tab <- demo %>% 
  select(-edvisit, -age) %>% 
  map(~{as_tibble(round(prop.table(xtabs(~ .x + edvisit, data = demo),2)*100,
                        1))}) %>% 
  bind_rows(.id = "var") %>% 
  mutate(edvisit = case_when(edvisit == 1 ~ 'yes_val',
                             edvisit == 0 ~ 'no_val')) %>% 
  spread(edvisit, n) %>% 
  filter(.x == 1) %>% 
  select(-.x)

# n table
n_tab <- demo %>% 
  select(-edvisit, -age) %>% 
  map(~{as_tibble(xtabs(~ .x + edvisit, data = demo))}) %>% 
  bind_rows(.id = "var") %>% 
  mutate(edvisit = case_when(edvisit == 1 ~ 'ed_yes_n',
                             edvisit == 0 ~ 'ed_no_n')) %>% 
  spread(edvisit, n) %>% 
  filter(.x == 1) %>% 
  select(-.x)

# demo_tab
demo_tab <- n_tab %>% 
  left_join(prec_tab, by = "var") %>% 
  bind_rows(age_tab, .)

knitr::kable(demo_tab, 
  caption = "Demographic characterstics by ED Visit (most vals are percent, expect age, which is mean")
```

Again, some missingness in variables. It would be good to have a conversaion on if we should impute these. I will probably need to transform the dummy variables to factor. 

## NO~2~ Distribution

Distribution of observed NO~2~ values.

```{r no2_dist}
no2_den_plot <- ggplot(asthma_no2_df, aes(x=no2)) +
  geom_density(color = '#9cecfb', fill = '#9cecfb', alpha = 0.7) +
  ylab("Density") +
  xlab(expression(paste("NO"[2], " ", mu,"g/m"^3))) +
  ryan_theme

# print no2 density plot
no2_den_plot
```

Assessing the relationship between NO~2~ and the proportion of ED visits by NO~2~ bins of 2 ug/m^3.

```{r no2_bin}
# checking linear relationship in p
no2_bin_df <- asthma_no2_df %>% 
  mutate(no2_bin = cut(no2, 
    breaks = seq(8, 24, by = 2), labels = seq(9, 23, by = 2))) %>% 
  group_by(no2_bin) %>% 
  summarise(n=n(), n_ed = sum(edvisit)) %>% 
  filter(!is.na(no2_bin)) %>% 
  mutate(ed_p = n_ed/n,
         no2 = as.numeric(as.character(no2_bin)))


# smooth spline fit
logit_smooth <- glm(edvisit ~ splines::ns(no2, 3), family = binomial(), 
                    asthma_no2_df) 

logit_lin <- glm(edvisit ~ no2, family = binomial(), asthma_no2_df) 

# predict based on no2 estimate
ns_ed <- predict(logit_smooth, data=asthma_no2_df, type = "response")
# ns_ed 95% CI around fit
ns_se <- predict(logit_smooth, data = asthma_no2_df, type = "response",
                      se.fit = T)$se.fit
# upper95 and lower95
ns_upper95 <- (ns_ed + 1.96*ns_se)
ns_lower95 <- (ns_ed - 1.96*ns_se)

# linear predictor
lin_ed <- predict(logit_lin, data = asthma_no2_df, type = "response")
# linear se
lin_se <- predict(logit_lin, data = asthma_no2_df, type = "response",
                      se.fit = T)$se.fit
# upper95 and lower95
lin_upper95 <- (lin_ed + 1.96*lin_se)
lin_lower95 <- (lin_ed - 1.96*lin_se)

# no2 vector
no2 <- asthma_no2_df$no2
# no2 ns spline
ed_ns_df <- data.frame(ns_ed, ns_lower95, ns_upper95, no2) %>% 
  rename(pr_ed = ns_ed, lower95 = ns_lower95, upper95 = ns_upper95) %>% 
  mutate(Fit = paste0("Spline (AIC: ", round(AIC(logit_smooth),1), ")"))

# lin fit
ed_lin_df <- data.frame(lin_ed, lin_lower95, lin_upper95, no2) %>% 
  rename(pr_ed = lin_ed, lower95 = lin_lower95, upper95 = lin_upper95) %>% 
  mutate(Fit = paste0("Linear (AIC: ", round(AIC(logit_lin),1), ")"))

# fit df
pr_ed_df <- bind_rows(ed_ns_df, ed_lin_df)

# plot of deciles by no2
fit_plot <- ggplot() +
  geom_point(data = no2_bin_df, aes(x = no2, y = ed_p), color = '#f2fcfe') +
  geom_line(data = pr_ed_df, aes(x=no2, y=pr_ed, group = Fit, color = Fit)) +
  geom_ribbon(data = pr_ed_df, aes(x=no2, ymin=lower95, ymax=upper95, 
                                   group = Fit, fill = Fit), alpha = 0.5) +
  scale_color_manual(values = c('#1c92d2', '#e100ff')) +
  scale_fill_manual(values = c('#1c92d2', '#e100ff')) +
  facet_wrap(~Fit) +
  ylab("Proportion of Emergency Department Visits") +
  xlab(expression(paste("Mean of NO"[2], " ", mu,"g/m"^3, " Bin"))) +
  ryan_theme +
  theme(legend.direction = "horizontal", legend.position = "bottom")

# print fit plot
fit_plot
```

Fit by AIC suggests that for a logit model, a linear fit is just as good a fit as splines I've tested (I tried 2 knots and 3 knots but only showing 3 knot spline). I think it's fine to leave continous NO2 as a linear predictor

## Potential Confounders

There are over 437 variables. I beleive Sheryl had identified confounders for the living < 500 meters of a roadway, but it's likely the potential confounders between NO~2~ cutoffs are different to the buffer, and perhaps even continous NO~2~. Individual schools are dummy-coded in so I'm going to take out the school id variable.

```{r possible_covs}
# subset to possible covariates based off sheryl's list
all_covs <- asthma_no2_df %>% 
  select(inbuff, male, enghome:olanghome, api:othrace, pubhouse,
         assthouse, hillres:estuary, pwhite:pinst, pmchhwoc:pimmb80nc,
         psamehome95:pworkmom, pmomnlf:urban) %>% 
  rename_all(funs(str_replace(., "_", ".")))

# nearzero variance of possible covs
nzv <- nearZeroVar(all_covs)
# filter to covariates that do not have near zero variance
covs_w_var <- all_covs[, -nzv]
# from 339 to 273

# create covariance matrix using non-parametric spearman to find vars with high
# correlation
corr_mat <- cor(covs_w_var, method = "spearman")

# find highly correlated predictors
high_cor <- findCorrelation(corr_mat, cutoff = 0.9)  

# filter to variables with lower correlation than 0.8
possible_covs <- covs_w_var[,-high_cor] 
# reduced to 273 to 225
```

### High-Density Propensity Score

I'm using the high-density propensity score algorigthm to further reduce the dimensions of covariates to only the 5 most influential. I like the high-density propensity score because it considers the relationship between both exposure and outcomes, and the possibility of that covariate for confounding. As we have many questionnaire covariates, environmental covariates, and census demographic covariates, I like this data reduction approach instead of building models piece by piece.

For simplicity, I'm going to fit the HDPS model based on the binary cutoff of NO~2~ at the upper decline of the NO~2~ distribution, which is a value of 19.9 ug/m^3^. It looks like this will work based on the fit plots below. It also gives us the counterfactual question:

What would the reported proportion of ED visits be in children with asthma if neighborhood NO~2~ were reduced below 19.9 ug/m^3^?

```{r hdps}
# trying out hdps to reduce to 50 vars
# step 2,3,4: identify covariates associated with treatment and outcome
hdps_vars <- hdps_screen(outcome = asthma_no2_df$edvisit, 
                         treatment = asthma_no2_df$no2_10dec,
                         covars = possible_covs, keep_n_per_dimension = 5,
                         keep_k_total = 10, verbose = T)

# output vector of covariates identified by HDS
hdps_vars_vec <- predict(hdps_vars) %>% 
  colnames() %>% 
  str_replace(., "_.*", "") %>% 
  unique()

# subset covariates to hdps_vars_vec
hdps_vars_df <- possible_covs[, colnames(possible_covs) %in% hdps_vars_vec]
# names
#print(colnames(hdps_vars_df))
```

The HDPS identified the following potential confounders: 

Subject-specific binary idicators of male, African American, and resident of a mixed housing land use area. 

Census-level variables percentage of children in census block? that speak Asian or Pacific Islander language, but not English well. Also, percentage American Indian, Native American of Latino population by census tract.

I will include these in adjusted models. Note, I realize that what might confounded the binary cutoff of NO~2~ may be different from continous NO~2~. I have not spent time identifying variables that may predict continous NO~2~. This is a limitation, but I think it's minor.  

## Standard Models

### Continous Relationship

Linear model fits these data just as well as a spline with 3 degrees of freedom.I'm going to estimate the linear relationship between a 1 ug/m^3^ increase in NO~2~ and the proportion/likelihood of reporting an asthma ED visit using both the identity link (difference in proportion) and the log link (relative difference in proportion).

```{r unadjusted_lin_relationships}
# unadjusted linear model
unadj_diff <- broom::tidy(glm(edvisit ~ no2, data=asthma_no2_df, 
             family = binomial(link="identity"))) %>% 
  filter(term == "no2") %>% 
  select(estimate, std.error) %>% 
  mutate(model = 'Logistic',
         type = "Difference",
         Y = "ED Visit",
         A = "NO2 continous",
         W = "Unadjusted",
         lower95 = round(estimate - 1.96*std.error,3),
         upper95 = round(estimate + 1.96*std.error,3),
         estimate = round(estimate, 3)) %>% 
  select(model, type, Y, A, W, estimate, lower95, upper95)

# relative
# unadjusted linear model
unadj_or <- broom::tidy(glm(edvisit ~ no2, data=asthma_no2_df, 
             family = binomial(link="log"))) %>% 
  filter(term == "no2") %>% 
  select(estimate, std.error) %>% 
  mutate(model = 'Logistic',
         type = "Odds Ratio",
         covariate = 'Unadjusted',
         Y = "ED Visit",
         A = "NO2 continous",
         W = "Unadjusted",
         lower95 = round(exp(estimate - 1.96*std.error),3),
         upper95 = round(exp(estimate + 1.96*std.error),3),
         estimate = round(exp(estimate), 3)) %>% 
  select(model, type, Y, A , W, estimate, lower95, upper95)

# table
knitr::kable(bind_rows(unadj_diff, unadj_or), 
  caption = 'Unadjusted linear relationship between NO2 and ED Visits')
```

Adjusted models adjusting for HDPS covariates. Note, I should really consider itteratiting for each model scenario.

```{r adj_linear_relationship}
# adjusted linear model
adj_diff <- broom::tidy(glm(edvisit ~ no2 + male + afra + mixhouse + 
                              pctlataian0 + kidasnengn, data=asthma_no2_df, 
             family = binomial(link="identity"))) %>% 
  filter(term == "no2") %>% 
  select(estimate, std.error) %>% 
  mutate(model = 'Logistic',
         type = "Difference",
         Y = "ED Visit",
         A = "NO2 continous",
         W = "HDPS",
         lower95 = round(estimate - 1.96*std.error,3),
         upper95 = round(estimate + 1.96*std.error,3),
         estimate = round(estimate, 3)) %>% 
  select(model, type, Y, A, W, estimate, lower95, upper95)

# adjusted linear model
adj_or <- broom::tidy(glm(edvisit ~ no2 + male + afra + mixhouse + 
            pctlataian0 + kidasnengn, data=asthma_no2_df, 
            family = binomial(link="log"))) %>% 
  filter(term == "no2") %>% 
  select(estimate, std.error) %>% 
  mutate(model = 'Logistic',
         type = "Odds Ratio",
         covariate = 'Unadjusted',
         Y = "ED Visit",
         A = "NO2 continous",
         W = "HDPS",
         lower95 = round(exp(estimate - 1.96*std.error),3),
         upper95 = round(exp(estimate + 1.96*std.error),3),
         estimate = round(exp(estimate), 3)) %>% 
  select(model, type, Y, A , W, estimate, lower95, upper95)

# table
knitr::kable(bind_rows(adj_diff, adj_or), 
  caption = 'Adjusted linear relationship between NO2 and ED Visits')
```

Adjusted variables differ very little from crude estimates.

### Binary Cutoff

Causal inference models like Targeted Maximum Likelihood (TMLE) generally require a binary point exposure, although there are exceptions when the relationship between A and Y is linear (g-estimation, structural equations modeling). I do believe there are extensions of TMLE now for a bounded continous A variable, but I have not really looked in to these methods yet.

```{r adj_linear_relationship_binary}
# adjusted linear model
adj_diff_binary <- broom::tidy(glm(edvisit ~ no2_10dec + male + afra + mixhouse + 
                              pctlataian0 + kidasnengn, data=asthma_no2_df, 
             family = binomial(link="identity"))) %>% 
  filter(term == "no2_10dec") %>% 
  select(estimate, std.error) %>% 
  mutate(model = 'Logistic',
         type = "Difference",
         Y = "ED Visit",
         A = "NO2 > 19.9",
         W = "HDPS",
         lower95 = round(estimate - 1.96*std.error,3),
         upper95 = round(estimate + 1.96*std.error,3),
         estimate = round(estimate, 3)) %>% 
  select(model, type, Y, A, W, estimate, lower95, upper95)

# unjusted linear model
unadj_or_binary <- broom::tidy(glm(edvisit ~ no2_10dec, data=asthma_no2_df, 
            family = binomial(link="log"))) %>% 
  filter(term == "no2_10dec") %>% 
  select(estimate, std.error) %>% 
  mutate(model = 'Logistic',
         type = "Odds Ratio",
         covariate = 'Unadjusted',
         Y = "ED Visit",
         A = "NO2 > 19.9",
         W = "Unadjusted",
         lower95 = round(exp(estimate - 1.96*std.error),3),
         upper95 = round(exp(estimate + 1.96*std.error),3),
         estimate = round(exp(estimate), 3)) %>% 
  select(model, type, Y, A , W, estimate, lower95, upper95)

# adjusted linear model
adj_or_binary <- broom::tidy(glm(edvisit ~ no2_10dec + male + afra + mixhouse + 
            pctlataian0 + kidasnengn, data=asthma_no2_df, 
            family = binomial(link="log"))) %>% 
  filter(term == "no2_10dec") %>% 
  select(estimate, std.error) %>% 
  mutate(model = 'Logistic',
         type = "Odds Ratio",
         covariate = 'HDPS',
         Y = "ED Visit",
         A = "NO2 > 19.9",
         W = "HDPS",
         lower95 = round(exp(estimate - 1.96*std.error),3),
         upper95 = round(exp(estimate + 1.96*std.error),3),
         estimate = round(exp(estimate), 3)) %>% 
  select(model, type, Y, A , W, estimate, lower95, upper95)

# table
knitr::kable(bind_rows(unadj_or_binary, adj_or_binary), 
  caption = 'Adjusted relationship between NO2 >19.9 and ED Visits')
```

## Targeted Maximum Likelihood Estimation

Using TMLE to estimate the marginal effect of reducing NO~2~ levels below 19.9 on reported ED visit.

Setting up counterfactual estimates of Q formula. The following code chunk is for collaborative TMLE, but I don't think I'll go that route since I've reduced the dimensions of possible covariates to 5 now.

```{r scalable_ctmle}
# lenght of observations
n <- length(asthma_no2_df$edvisit)
# observed Y
Y <- asthma_no2_df$edvisit
# observed X
A <- asthma_no2_df$no2_10dec
# covariates
W <- as.matrix(hdps_vars_df)
# start by building initial estimate of Q
# counterfactual estimate of Y given all exposure is 0
Ya0 <- rep(mean(Y[A == 0]),n)
# counterfactual estimate of Y given all exposure is 1
Ya1 <- rep(mean(Y[A == 1]),n)
# assign initial estimate of Q e
Q <- cbind(Ya0, Ya1)
```


Running a standard TMLE model.

```{r tmle_model}
# sl lib
sl_lib <- c("SL.glmnet", "SL.glm", "SL.step")
# tmle model
tmle_mod <- tmle(Y = Y, A = A, W = W, 
                 Q.SL.library = sl_lib, g.SL.library = sl_lib, 
                 family = "binomial", V = 10)
```

Checking propensity of NO~2~ exposure from TMLE model for ETA violation. Looks like we are okay.

```{r tmle_g_eta}
# predict on test
g_pred <- as_data_frame(tmle_mod$g$g1W) %>% 
  rename(g = value)

# histogram of predicted values
eta_plot <- ggplot(data.frame(g_pred), aes(x=g)) +
  geom_density(fill = "#bdfff3", color = "#bdfff3", alpha = 0.5) +
  xlim(0, 1) +
  ylab('Density') +
  xlab(expression(paste("Probability of NO"[2], " >19.9 ", mu,"g/m"^3, " | W"))) +
  ryan_theme
# print eta plot
eta_plot
```

TMLE Estimates.

```{r tmle_est}
tmle_mod
```


## Collaborative TMLE

I've run this, but it won't improve on TMLE since I've reduce variables to only 5 and made sure they balance the propensity. Consider coming back to this with more dimensions for the paper.

Building initial lasso fit of g formula.

```{r g_form}
# lasso fit
lasso_fit <- cv.glmnet(y=A, x=W, family = "binomial", nlambda =100, 
                       nfolds = 5)
```

Build sequence of lmabdas from lambda selected by CV.
```{r lasso_lambdas}
lasso_lambdas <- lasso_fit$lambda[lasso_fit$lambda <=
                                  lasso_fit$lambda.min][1:10]
```

Custom lasso superlearner library.

```{r sl_lasso_template}
# Build SL template for glmnet
SL.glmnet_new <- function(Y, X, newX, family, obsWeights, id, alpha = 1,
                           nlambda = 100, lambda = 0,...){
      # browser()
      if (!is.matrix(X)) {
            X <- model.matrix(~-1 + ., X)
            newX <- model.matrix(~-1 + ., newX)
      }
      fit <- glmnet::glmnet(x = X, y = Y,
                            lambda = lambda,
                            family = family$family, alpha = alpha)
      pred <- predict(fit, newx = newX, type = "response")
      fit <- list(object = fit)
      class(fit) <- "SL.glmnet"
      out <- list(pred = pred, fit = fit)
      return(out)
}

# Use a sequence of estimator to build gn sequence:
SL.cv1lasso <- function (... , alpha = 1, lambda = lasso_lambdas[1]){
      SL.glmnet_new(... , alpha = alpha, lambda = lambda)
}

SL.cv2lasso <- function (... , alpha = 1, lambda = lasso_lambdas[2]){
      SL.glmnet_new(... , alpha = alpha, lambda = lambda)
}

SL.cv3lasso <- function (... , alpha = 1, lambda = lasso_lambdas[3]){
      SL.glmnet_new(... , alpha = alpha, lambda = lambda)
}

SL.cv4lasso <- function (... , alpha = 1, lambda = lasso_lambdas[4]){
      SL.glmnet_new(... , alpha = alpha, lambda = lambda)
}

SL.library = c('SL.cv1lasso', 'SL.cv2lasso', 'SL.cv3lasso', 'SL.cv4lasso', 'SL.glm')
```

Construct the object folds, which is a list of indices for each fold.

```{r lasso_folds}
V = 5
folds <- by(sample(1:n,n), rep(1:V, length=n), list)
```

Use folds and SuperLearner template to compute gn_candidates and gn_candidates_cv.

```{r gn_seq}
gn_seq <- build_gn_seq(A = A, W = W, SL.library = SL.library, folds = folds)
```

C-TMLE only has difference estiamtes right now. I looked at the function and believe I can modify it if we decided to estimate the ratio.

```{r ctmle_fit}
# run ctmle model
ctmle_general_fit1 <- ctmleGeneral(Y = Y, A = A, W = W, Q = Q,
                                   ctmletype = 1, family = "binomial",
                                   gn_candidates = gn_seq$gn_candidates,
                                   gn_candidates_cv = gn_seq$gn_candidates_cv,
                                   folds = folds, V = 5)
# print fit
ctmle_general_fit1
```

## Comparison of Binary Estimates Using Standard Model vs. TMLE

```{r results_tab}
# tmle_results
tmle_or <- round(tmle_mod$estimates$OR$psi,3)
tmle_lower95 <- round(tmle_mod$estimates$OR$CI, 3)[1]
tmle_upper95 <- round(tmle_mod$estimates$OR$CI, 3)[2]

# tmle_results 
tmle_results <- data.frame(tmle_or, tmle_lower95, tmle_upper95) %>% 
  rename(estimate = tmle_or, lower95 = tmle_lower95, upper95 = tmle_upper95) %>% 
  mutate(model = "TMLE", type = "Odds Ratio", Y = "ED Visit", A = "NO2 > 19.9",
         W = "HDPS")

relative_results <- bind_rows(unadj_or_binary, adj_or_binary, tmle_results) %>% 
  mutate(Model = paste0(model, ": ", W))

# table
knitr::kable(relative_results, 
  caption = 'Results for different models between NO2 >19.9 and ED Visits')
```

Plot of same results.

```{r results_plot}
# results plot
results_plot <- ggplot(data = relative_results, aes(x = Model, y = estimate,
                                                    color = model, group = model)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.3) +
  scale_color_manual("Model Type", values = c("#9cecfb", "#ff00cc")) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "red") +
  ylim(0.8, 4) +
  ylab(expression(paste("Odds Ratio: NO"[2], " > 19.9 ", mu,"g/m"^3, 
                        " (Yes vs. No)"))) +
  xlab("Model") +
  ryan_theme +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted"),
        panel.grid.minor = element_blank(),
        legend.direction = "horizontal", legend.position = "bottom")
# print results
results_plot
```

## Conclusion

We implemented a standard and causal inference framework to assess the relationship between NO2 and ED visits and found strengths and limitations to both approaches. The standard modeling approach allowed us to evaluate the concentration-response relationship between the exposure and outcome, but these results are harder to translate in to actionable policy or public health decisions. In contrast, results from our TMLE formalizes our counterfactual question on the population effect if we reduced NO2 below 19.9 µg/m^3^. However, this association cannot be interpreted causally due to violations in conditional exchangeability and the cross-sectional nature of NO~2~ assessment and reported ED visits.

I just copied this statement from the abstract. I will write some more stuff.