---
title: "NO2 and asthma using TMLE"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Random Forest to find out covariates that predict binary NO2.

```{r packages}
library(tidyverse)
library(randomForest)
library(tmle)
library(SuperLearner)
```

Loading Oakland Kicks Asthma dataframe and joining with NO~2~ dataframe.
```{r read_data}
no2layer <- haven::read_sas("./no2layer.sas7bdat") %>% # summary of NO2
  select(OKAID, NO2) %>% 
  mutate(OKAID = as.character(OKAID))
  
no2_col_names <- colnames(no2layer) %>% tolower() # lowercase all col names
colnames(no2layer) <- no2_col_names # assign lowercase col names to df

#head(no2layer) # check out the first 6 lines
#glimpse(no2layer)

okawork <- haven::read_sas("./okawork1.sas7bdat") %>% 
  # convert okaid to character
  mutate(okaid = as.character(okaid))

# join okawork with no2layer by okaid
asthma_no2_df <- okawork %>% right_join(no2layer, by = "okaid") %>% 
  # create ed visit variable from sh_11 var (assuming 1=event)
  mutate(edvisit = ifelse(sh_11 == 1, 1,
                   ifelse(sh_11 == 2, 0, NA))) %>% 
  # filter out missing edvisits
  filter(!is.na(edvisit)) %>%
  # filter to prevalent asthma cases
  # select(edvisit, asthma, inbuff, mfi_t, adengonly, eslgood, piprlt50, pctlata,
  #        ph75_100k, pctutilinc, pctotown, pctnofuel, kidspanengg, pkgt6sp1,
  #        regcomm, commcomr, urbanres, no2) %>% 
  filter(asthma == 1) %>% 
  mutate(no2_cut = ifelse(no2 >= 19, 1, 0),
         no2_median = ifelse( no2 >= 17.197, 1, 0))
```

There are over 437 variables. I beleive Sheryl had identified confounders for the living <500 meters of a roadway, but it's likely the potential confounders are different

```{r possible_covs}
possible_covs <- asthma_no2_df %>% 
  select(school, grade, male, enghome:olanghome, api:othrace, pubhouse, assthouse,
         hillres:estuary, pwhite:pinst, pmchhwoc:pimmb80nc, psamehome95:pworkmom,
         pmomnlf:urban)
# no2 cut
no2_cut <- asthma_no2_df$no2_cut
# asthma ed visit
edvist <- asthma_no2_df$edvisit
```

Random forest to predict NO~2~ cut.

```{r no2_rf}
# no2 randomforest model
no2_rf <- randomForest(as.factor(no2_cut) ~ ., data = possible_covs,
                       importance=T, ntree=500)

varImpPlot(no2_rf)

# view a tree
reprtree:::plot.getTree(no2_rf)

pred <- predict(no2_rf, possible_covs)

xtabs(~no2_cut + pred)
```

```{r asthma_rf}
# asthma randomforest model
edvist_rf <- randomForest(as.factor(edvist) ~ ., data = possible_covs,
                       importance=T, ntree=500)

varImpPlot(edvisit_rf)

# view a tree
reprtree:::plot.getTree(no2_rf)

pred <- predict(no2_rf, possible_covs)

xtabs(~no2_cut + pred)
```