---
title: "Oakland Kicks Asthma MSM"
author: "Ryan_Gan"
date: "April 25, 2017"
output: html_document
---

## Marginal Structural Model

General purpose of this project is to try and estimate an MSM that doesn't have an ETA violation.

## Research Question

Aim: Utilize a MSM to determine the causal relation between residential proximity to freeway and asthma-related morbidity in children with asthma.

Study population: Cohort of school children with asthma living in Oakland, California.
Outcome: Emergency room visit.
Exposure: Living within 500 meters of a freeway.

## Replication of Sheryl's Dissertation Results

Importing SAS files of Oakland Kicks Asthma data provided by Sheryl.

```{r library setup, echo = F, warning = F, message = F}
library(haven) # library for importing SAS files
library(tidyverse) # data wrangle package
library(broom)
library(tmle) # targeted maximum likelihood package
library(SuperLearner) # ensemble method package
#library(nnet) # neural net algorithm package
#library(randomForest) # random forest algorithm package
```

Importing SAS data set and limiting to covariates found in Sheryl's code. I could not find some covariates, so I used some variables that were similar. Example is regcomm. I am also going to filter prevalent asthma, indicated by asthma = 1.

```{r import and data wrangle, echo = F, warning = F, message = F}
no2layer <- read_sas("./no2layer.sas7bdat") %>% # summary of NO2
  select(OKAID, NO2) %>% 
  mutate(OKAID = as.character(OKAID))
  
no2_col_names <- colnames(no2layer) %>% tolower() # lowercase all col names
colnames(no2layer) <- no2_col_names # assign lowercase col names to df

#head(no2layer) # check out the first 6 lines
#glimpse(no2layer)

okawork <- read_sas("./okawork1.sas7bdat") %>% 
  # convert okaid to character
  mutate(okaid = as.character(okaid))
  
# join okawork with no2layer by okaid
asthma_df <- okawork %>% right_join(no2layer, by = "okaid") %>% 
  # create ed visit variable from sh_11 var (assuming 1=event)
  mutate(edvisit = ifelse(sh_11 == 1, 1,
                   ifelse(sh_11 == 2, 0, NA))) %>% 
  # filter out missing edvisits
  filter(!is.na(edvisit)) %>%
  # filter to prevalent asthma cases
  select(edvisit, asthma, inbuff, mfi_t, adengonly, eslgood, piprlt50, pctlata,
         ph75_100k, pctutilinc, pctotown, pctnofuel, kidspanengg, pkgt6sp1,
         regcomm, commcomr, urbanres) %>% 
  filter(asthma == 1)
# in Sheryl's SAS code, she has regcomm_perc but I could not find it, so using
# regcomm for now.

```

2x2 table of ED visit by living within 500 m buffer.

```{r descriptives, echo = F, warning = F, message = F}
# edvist by inbuff in children with asthma
xtabs(~ edvisit + inbuff, data = asthma_df)
# caluclated crue odds ratio
(20/178)/(48/550)
```

Crude odds ratio calculations using standard logistic regression.

```{r prev asthma standard estimates, echo = F, warning = F, message = F}
# prevalent asthma 
# crude model
crude_mod <- tidy(glm(edvisit ~ inbuff, data = asthma_df,
                    family = "binomial"(link = "logit")))

# crude odds ratio and 95%CI
vals <- round(exp(c(crude_mod[2,2], crude_mod[2,2]-1.96*crude_mod[2,3], 
      crude_mod[2,2]+1.96*crude_mod[2,3])),2)
rbind(c("OR", "Lower_95", "Upper_95"), vals)
```

Adjusted logistic model using covariates that will go in to SuperLearner. Covariates were selected during Sheryl's dissertation.

```{r prev asthma adjusted model, echo = F, warning = F, message = F}
# adjusted model
adj_mod <- tidy(glm(edvisit ~ inbuff + mfi_t + adengonly + eslgood + 
          piprlt50 + pctlata + ph75_100k + pctutilinc + pctotown + pctnofuel + 
          kidspanengg + pkgt6sp1 + regcomm + commcomr + urbanres, 
          data = asthma_df, family = "binomial"(link="logit")))

# adjusted odds ratio and 95%CI
vals <- round(exp(c(adj_mod[2,2], adj_mod[2,2]-1.96*adj_mod[2,3], 
      adj_mod[2,2]+1.96*adj_mod[2,3])),2)
rbind(c("OR", "Lower_95", "Upper_95"), vals)
```

### IPTW

IPTW function (custom function).

```{r eta check, echo = F, warning = F, message = F}
a_mod <- glm(inbuff ~ mfi_t+adengonly+eslgood+piprlt50+pctlata+ph75_100k+
  pctutilinc+pctotown+pctnofuel+kidspanengg+pkgt6sp1+commcomr+urbanres, 
  data = asthma_df, family = "binomial")

# estimate probability of a 
pr_a <- predict(a_mod, type = "response")

ggplot(as.tibble(pr_a), aes(value)) +
  geom_density()
```

Looks like we have a possible ETA violation.The distribution of living with 500 meters of a major road-way appears to be almost perfectly predicted by some of our covariates.

Running IPTW with bootstrap to see results.

```{r iptw boot, echo = F, warning = F, message = F}
# percentile method with broom
set.seed(321)
boot_iptw <- asthma_df %>% bootstrap(1000) %>% 
  do(tidy(MSModelR::iptw(data=., y="edvisit", a="inbuff", w=paste0("mfi_t+adengonly",
  "+eslgood+piprlt50+pctlata+ph75_100k+pctutilinc+pctotown+pctnofuel+",
  "kidspanengg+pkgt6sp1+commcomr+urbanres"), estimate="ratio")))

# create empty matrix
estimates <- matrix(nrow = 1, ncol = 3, byrow = T)
colnames(estimates) <- c("boot_median", "lower_95", "upper_95")

# fill matrix 
estimates[,1] <- exp(round(quantile(boot_iptw$x, 0.5),3))
estimates[,2] <- exp(round(quantile(boot_iptw$x, 0.025),3))
estimates[,3] <- exp(round(quantile(boot_iptw$x, 0.975),3))

# convert matrix to dataframe
est_df <- data.frame(estimates)

knitr::kable(est_df, caption = paste0("IPTW estimates with bootstrapped 95%CI"))
```

Percentile bootstrap method produces really high upper bound for CI with this data set. Could be because possible IPTW violation.

We'll try with g-formula (simple substitution) to see if we get different results.
```{r gform, echo=F, message=F, warning=F}
# set seed
set.seed(123)
boot_gform <- asthma_df %>% bootstrap(1000) %>% 
  do(tidy(MSModelR::g.formula(data=., y="edvisit", a="inbuff", 
  q.model=paste0("inbuff+mfi_t+adengonly+eslgood+piprlt50+pctlata+ph75_100k+",
  "pctutilinc+pctotown+pctnofuel+kidspanengg+pkgt6sp1+commcomr+urbanres"), 
  model.family = "binomial", estimate="ratio")))

# create empty matrix
estimates <- matrix(nrow = 1, ncol = 3, byrow = T)
colnames(estimates) <- c("boot_median", "lower_95", "upper_95")

# fill matrix 
estimates[,1] <- exp(round(quantile(boot_gform$x, 0.5),3))
estimates[,2] <- exp(round(quantile(boot_gform$x, 0.025),3))
estimates[,3] <- exp(round(quantile(boot_gform$x, 0.975),3))

# convert matrix to dataframe
est_df <- data.frame(estimates)

knitr::kable(est_df, caption = paste0("G-formula estimates with bootstrapped 95%CI"))
```

G-formula produces estimates that are very different from IPTW and our standard regression estimate method. Not quite sure why the estimates are so different... I'll need to think about it.

### TMLE

Targeted maximum likelihood (TMLE) uses an ensemble of machine learning algorithms to optimize the covaraites that predict treatment and the q-model. It then uses a doubly-robust method to estimate the causal estimand, Not really sure what covariates to feed to the models, but I'm going to just mess around with TMLE for now with the data frames I have. 

Also, I can feed every variable other than treatment and the outcome in to the model and let it decide what's the best fit.

```{r tmle msm}
# defining superlearner libraries to use
# since i'm not crossvalidating yet
sl_lib <- c("SL.glm","SL.step","SL.glm.interaction")


tmle_mod <- tmle(Y = asthma_df$edvisit, A = asthma_df$inbuff, 
                 W = subset(asthma_df, select = -c(edvisit, inbuff, asthma)),
                 Q.SL.library = sl_lib, g.SL.library = sl_lib, 
                 V = 10, family = "binomial")
tmle_mod
```

I get similar answers to the crude and adjusted as IPTW, but confidence intervals are much tighter.

## Next steps

Make sure the machine-learning steps in the TMLE model is accurate (run cross-validations and sensitivity tests).

Look at the exposure, NO~2~. We will have to use a binary classification (NO~2~ cut point) or some comparisons of ordinal variables (high vs low). This will result in a loss of information, but may be more relevant to public health.

Try structural nested models using DTreg package,e. SNMs are MSMs that allow continuous variables. In the past, they were difficult to use and implement, thus were not widely used. This is appealing because I think we can answer questions like "what if we brought NO~2~ down to the mean level of ...?".

```{r adding no2, echo=F, warning=F, message=F}
# join okawork with no2layer by okaid
asthma_no2_df <- okawork %>% right_join(no2layer, by = "okaid") %>% 
  # create ed visit variable from sh_11 var (assuming 1=event)
  mutate(edvisit = ifelse(sh_11 == 1, 1,
                   ifelse(sh_11 == 2, 0, NA))) %>% 
  # filter out missing edvisits
  filter(!is.na(edvisit)) %>%
  # filter to prevalent asthma cases
  select(edvisit, asthma, inbuff, mfi_t, adengonly, eslgood, piprlt50, pctlata,
         ph75_100k, pctutilinc, pctotown, pctnofuel, kidspanengg, pkgt6sp1,
         regcomm, commcomr, urbanres, no2) %>% 
  filter(asthma == 1) %>% 
  mutate(no2_cut = ifelse(no2 >= 19, 1, 0))
```

Distribution of NO~2~ values and some summary stats.
```{r no2 plot, echo=F}
no2_dist_plot <- ggplot(data = asthma_no2_df, aes(x=no2)) +
  geom_density()

no2_dist_plot

summary(asthma_no2_df$no2)
```

Crude association between increasing NO~2~ and ED visits in a population with asthma.
```{r no2 unadjusted}
no2_mod <- tidy(glm(edvisit ~ no2, family = "binomial", data = asthma_no2_df))

# crude odds ratio and 95%CI
vals <- round(exp(c(no2_mod[2,2], no2_mod[2,2]-1.96*no2_mod[2,3], 
      no2_mod[2,2]+1.96*no2_mod[2,3])),2)
rbind(c("OR", "Lower_95", "Upper_95"), vals)
```

Mean level of NO~2~ by ED visit group.
```{r mean no2, echo = F}
mean_no2 <- tidy(glm(no2 ~ edvisit, family = "gaussian", data = asthma_no2_df))
mean_no2
```

Adjusting for the same covariates we adjusted for in our buffer analyses.
```{r no2 adjusted}
no2_mod <- tidy(glm(edvisit ~ no2+mfi_t+adengonly+eslgood+piprlt50+pctlata+
  ph75_100k+pctutilinc+pctotown+pctnofuel+kidspanengg+pkgt6sp1+commcomr+urbanres, 
  family = "binomial", data = asthma_no2_df))

vals <- round(exp(c(no2_mod[2,2], no2_mod[2,2]-1.96*no2_mod[2,3], 
      no2_mod[2,2]+1.96*no2_mod[2,3])),2)
rbind(c("OR", "Lower_95", "Upper_95"), vals)
```

We might have overadjusted. Signal is still relatively strong, but our 95% CIs now contain 1. I plan to actually assess confounding criteria and build models with DAGs, just being a little lazy here...

```{r linearity assumption, echo = F}
# prevalence by bin
no2_bins <- seq(from = 6, to = 24, by = 2)
no2_labels <- seq(from = 7, to = 23, by = 2)

no2_prev_bin <- asthma_no2_df %>% 
  mutate(no2_bin = cut(no2, breaks = no2_bins, labels = no2_labels)) %>% 
  group_by(no2_bin) %>% 
  summarise(n = n(), ed_sum = sum(edvisit)) %>% 
  mutate(ed_prev = ed_sum/n, no2 = as.numeric(as.character(no2_bin))) %>% 
  select(-no2_bin)
# plot
ggplot(data=asthma_no2_df, aes(x=no2, y=edvisit)) +
  geom_smooth(method = "loess", se=T) + 
  geom_smooth(method = "lm", formula=y~x, se=F, color = "red") +
  geom_point(data = no2_prev_bin, aes(x=no2, y=ed_prev)) +
  theme_bw()
```

I think the relationship between NO~2~ and ED visits can be more or less represented by a linear relationship. Smoothed function is a little funky. 

Anyways, let's create a cut point at 19, where the point kind of jumps to try out some of our MSM approaches.

First some standard calculations.
```{r 2x2 no2 cut}
xtabs(~no2_cut + edvisit, data= asthma_no2_df)
(60/151)/(138/447)

no2_mod <- tidy(glm(edvisit ~ no2_cut, family = "binomial", data = asthma_no2_df))

# crude odds ratio and 95%CI
crude_vals <- round(exp(c(no2_mod[2,2], no2_mod[2,2]-1.96*no2_mod[2,3], 
      no2_mod[2,2]+1.96*no2_mod[2,3])),2)
rbind(c("OR", "Lower_95", "Upper_95"), crude_vals)

no2_adj_mod <- tidy(glm(edvisit ~ no2_cut+mfi_t+adengonly+eslgood+piprlt50+pctlata+
  ph75_100k+pctutilinc+pctotown+pctnofuel+kidspanengg+pkgt6sp1+commcomr+urbanres, 
  family = "binomial", data = asthma_no2_df))

adj_vals <- round(exp(c(no2_adj_mod[2,2], no2_adj_mod[2,2]-1.96*no2_adj_mod[2,3], 
      no2_adj_mod[2,2]+1.96*no2_adj_mod[2,3])),2)
rbind(c("OR", "Lower_95", "Upper_95"), adj_vals)
```

MSM approach.

```{r no2 iptw boot, echo=F}
# percentile method with broom
set.seed(321)
boot_iptw <- asthma_no2_df %>% bootstrap(1000) %>% 
  do(tidy(MSModelR::iptw(data=., y="edvisit", a="no2_cut", 
  w=paste0("mfi_t+adengonly+eslgood+piprlt50+pctlata+ph75_100k+pctutilinc+",
  "pctotown+pctnofuel+kidspanengg+pkgt6sp1+commcomr+urbanres"), 
  estimate="ratio")))

# create empty matrix
estimates <- matrix(nrow = 1, ncol = 3, byrow = T)
colnames(estimates) <- c("boot_median", "lower_95", "upper_95")

# fill matrix 
estimates[,1] <- exp(round(quantile(boot_iptw$x, 0.5),3))
estimates[,2] <- exp(round(quantile(boot_iptw$x, 0.025),3))
estimates[,3] <- exp(round(quantile(boot_iptw$x, 0.975),3))

# convert matrix to dataframe
est_df <- data.frame(estimates)

knitr::kable(est_df, caption = paste0("IPTW estimates with bootstrapped 95%CI"))
```

Definately hit some ETA vilations for some of the boot iterations. 

```{r tmle no2 cut}
sl_lib <- c("SL.glm","SL.step","SL.glm.interaction")

tmle_mod <- tmle(Y = asthma_no2_df$edvisit, A = asthma_no2_df$no2_cut, 
                 W = subset(asthma_no2_df, 
                            select = -c(edvisit, inbuff, asthma,no2_cut,no2)),
                 Q.SL.library = sl_lib, g.SL.library = sl_lib, 
                 V = 10, family = "binomial")
tmle_mod
```

